CODE CHECK

### 2020-09-15 codecheck-paper.Rmd
rough edits from discussion

* Prefer longlines or short?
* Can we switch to latex for this?  (Find it slightly easier.)
* Which journal to target -- maybe worry abotu that once we have a
  preprint.  F1000R my current favourite?

* Check with Giuliano re: "code works".

@@

So, assuming that researchers begin to share more artifacts, how
might they be examined to test that they do what they claim?  For
example, most scientific journals now require a data sharing statement
that outlines what data the authors have (or will) share.  The
implementation status of this varies considerably according to the
journal.  At one of the spectrum, there are specialist journals  that
have been created to accept "data papers" (e.g. *Scientific Data*,
... NAME COUPLE OF OTHERS).  These journals have established rigorous
procedures by which the data is validated according to standards in
each field.  At the other end of the spectrum, we still witness today
the infamous statement "Data available upon reasonable request"
whereby authors, well-intentioned, often cannot provide the data when
readers ask for it.

@@

in house staff, pool of ECRs, volunteeres,

@@

**3. Credit is given to codecheckers.** (TODO Explain in a few
sentences what each means, e.g. see 1,2)

@@

@@

1. Gigascience -- our first paper.  Only visualisation, not machine
   learning (which would have taken sseveral days of compute time).
2. report 9 -- covid model of UK lockdown model.
3. Geoscience

@@

@@

To run the system, we rely on github to host our work, and use a
custom R pacakge, codecheck, to help with the generation of
certificates and depositing work on zenodo.  (Cite zen4R).

@@

@@

@@

Handling failure:  we have no established process for this, as so far
all our codechecks succeeded.  Open question remains as what to do in
this case -- should we publicly report that the code would not run.

Compute time:  for those papers that take significant compute time
(think days, not minutes), who will pay for the compute time?

Software: authors can currently provide code that requires
propieratiry software.  Given the prevalence of software like MATLAB,
this seems like a pragmatic choice, but then (a) we can't rely on
open infrastructure like mybinder, (b) codechecker must have that
software.

Can't someone cheat? yes.  we don't check it is correct code, just
simply that it runs.  Very low bar.  But by having the code open,
interested parties can then examine the code, and knowing that the
code will be open we hope will be an incentive for authors to share.

Who's got time for more peer review?  Agree, but ECRS might be
attracted to working with groups (recall it is not anonymous)

Who finds the codechecker?  Good point!  The journal.

@@

- How to embed this in juornals.
- ???

>>>>

* Prefer one sentence per line / short / one para per line.
* Nust* and Eglen*. contributed equally.
* latex.
* Which journal to target -- maybe worry abotu that once we have a
  preprint.  F1000R my current favourite.

* Check with Giuliano re: "code works".  Write to Naomi to apologise / placeholder.

@@

So, assuming that researchers begin to share more artifacts, how might they be examined to test that they do what they claim?  For example, most scientific journals now require a data sharing statement that outlines what data the authors have (or will) share.  The implementation status of this varies considerably according to the journal.  At one of the spectrum, there are specialist journals that have been created to accept "data papers" (e.g. *Scientific Data*, ... NAME COUPLE OF OTHERS).  These journals have established rigorous procedures by which the data is validated according to standards in each field.  At the other end of the spectrum, we still witness today the infamous statement "Data available upon reasonable request" whereby authors, well-intentioned, often cannot provide the data when readers ask for it.
(ANDREW H: any references to sum up poor/evolving state of data sharing; Vines et al 2014).

@@

in house staff, pool of ECRs, volunteeres,
for inhouse staff, credit not an issue.

@@

**3. Credit is given to codecheckers.** (TODO Explain in a few sentences
what each means, e.g. see 1,2) Ideally publons, or similar, but can just
be a citation to the publicly available DOI.  (Community DOIs use
these.)

@@

Do we want to list them in priority?  Open by default.  Community principles.

@@

1. Gigascience -- our first paper.  Only visualisation, not machine
   learning (which would have taken sseveral days of compute time).
2. report 9 -- covid model of UK lockdown model.
3. Geoscience  (add 6 for Agile).

@@

What can you expect to find a certificate?  Draw an annotated example.

@@

### Tools
To run the system, we rely on github to host our work, and use a
custom R pacakge, codecheck, to help with the generation of
certificates and depositing work on zenodo.  (Cite zen4R).

You can do all with open infrastructure (e.g. automated rendering of
register). but you can use your own tools if you prefer.

@@

TO BE CONTINUED.

@@

DELIBERATEY NOT COMPREHENSIVE HERE.

@@

**Handling failure**:  we have no established process for this, as so far
all our codechecks succeeded.  Open question remains as what to do in
this case -- should we publicly report that the code would not run.

**Compute time**:  for those papers that take significant compute time
(think days, not minutes), who will pay for the compute time?

**Software**: authors can currently provide code that requires
propieratiry software.  Given the prevalence of software like MATLAB,
this seems like a pragmatic choice, but then (a) we can't rely on
open infrastructure like mybinder, (b) codechecker must have that
software.

**Can't someone cheat?** yes.  we don't check it is correct code, just
simply that it runs.  Very low bar.  But by having the code open,
interested parties can then examine the code, and knowing that the
code will be open we hope will be an incentive for authors to share.

**Who's got time for more peer review?**  Agree, but ECRS might be
attracted to working with groups (recall it is not anonymous)

**Who finds the codechecker?** Good point!  The journal.  Community
however think this is an important issue, and are volunteering.

@@

- How to embed this in juornals.
- When is CODECHECK redundant?  will it become part of regular peer review?
or, if ayuthors become more skillful, then everything becomes automated, and no need for codecheck.
CODECHECK disappearing is then a measure of success.

Or will it always be adjunct to peer review, and needed?

working within the trad system of publishers to help push the conversation along.
